---
title: "ModelingAssignments"
author: "Zoe Li"
date: '`r Sys.Date()`'
output: pdf_document
site: bookdown::bookdown_site
editor_options: 
  markdown: 
    wrap: 72
---

# Modeling

# Exercises

1.  Given a function (relationship between two variables): $$
    y=\frac{5 x^{3}}{4 x+3}
    $$ Find $\frac{d y}{d x}$ (the first derivative or slope of the
    function).

Answer

```{r}
f=expression((5*x^3)/(4*x+3))
```

```{r}
D(f, 'x')
```

2.  Find the slope of the following functions:

<!-- -->

(a) $y=(5 x-7)^{3}$
(b) $y=\left(x^{2}+3 x-1\right)^{5}$

Answer

```{r}
f=expression((5*x-7)^3)
```

```{r}
D(f, 'x')
```

```{r}
f=expression((x^2+3*x-1)^5)
```

```{r}
D(f, 'x')
```

3.  Find the first and second derivatives of the following functions:

<!-- -->

(a) $y=e^{5 x}$
(b) $y=a^{x}$
(c) $y=\ln 2 x$

Answer

```{r}
f=expression(e^(5*x))
```

```{r}
D(f, 'x')
```

```{r}
D(D(f,'x'),'x')
```

```{r}
f=expression(a^x)
```

```{r}
D(f, 'x')
```

```{r}
D(D(f,'x'),'x')
```

```{r}
f=expression(log(exp(2*x)))
```

```{r}
D(f, 'x')
```

```{r}
D(D(f,'x'),'x')
```

4.  Given the demand function $Q=650-5 P-P^{2}$, where $Q$ denotes
    quantity sold and $P$ denotes the price, find the price elasticity
    of demand when $P=10$.

Answer A good's price elasticity of demand ({\displaystyle
E\_{d}}E\_{d}, PED) is a measure of how sensitive the quantity demanded
is to its price.

It may also be defined as the ratio of the percentage change in quantity
demanded to the percentage change in price of particular commodity.

$$
E_{\langle p\rangle}=\frac{\Delta Q / Q}{\Delta P / P}
$$

```{r}
Q=expression(Q=650-5*P-P^2)
```

The first derivative with refer to P, or dQ/dP

```{r}
D(Q, 'P')
```

The price elasticity of demand is (-(5 + 2 \* P))\*(P/Q)

Since Q=650-5\*P-P\^2

When P is equal to 10, the price elasticity of demand is

```{r}
(-(5 + 2 * 10))*((650-5*10-10^2)/10)
```

5.  Given that $TC=31+24 Q-5.5 Q^{2}+\frac{1}{3} Q^{3}$, where $T C$
    denotes total cost and $Q$ denotes quantity sold, find the values of
    $Q$ where the cost function reaches a minimum and a maximum. How can
    you verify if it is a minimum or a maximum? Find the total cost at
    the minimum and maximum points.

Answer

```{r}
TC=expression(31+24*Q-5.5*Q^(2)+(1/3)*Q^(3))
```

The first derivative

```{r}
D(TC, 'Q')
```

The simplized form

24-11\*Q+Q\^2

When the first derivative is equal to 0:

Q=3 or 8

The second derivative

```{r}
D(D(TC,'Q'),'Q')
```

The simplified form

-11+2Q

When Q = 3:

the second derivative (=-5) is less than 0;

When Q = 8:

the second derivative (=5) is greater than 0.

According to the rule: The second derivative is positive (f (x) \> 0):
When the second derivative is positive, the function f(x) is concave up.
2. The second derivative is negative (f (x) \< 0): When the second
derivative is negative, the function f(x) is concave down.

When Q = 3, the total cost should be the maximum, which is

```{r}
31+24*3-5.5*3^(2)+(1/3)*3^(3)
```

When Q = 8, the total cost should be the minimum, which is

```{r}
31+24*8-5.5*8^(2)+(1/3)*8^(3)
```

6.  A monopolist sells two products, $x$ and $y$, for which the demand
    functions are: $$
    \begin{aligned}
    &Q_{x}=25-0.5 P_{x} \\
    &Q_{y}=30-P_{y}
    \end{aligned}
    $$ and the combined cost function is $$
    c=Q_{x}^{2}+2 Q_{x} Q_{y}+Q_{y}^{2}+20
    $$ Find the following:

<!-- -->

(a) the profit-maximizing level of output for each product
(b) the profit-maximizing price for each product, and
(c) the maximum profit.

Answer

Since profit=sales revenue-cost =Q\*P-c,

in this case, profit=Px*Qx+Py*Qy-cost
=Qx*(-2Qx+50)+Qy*(-Qy+30)-(Qx^2+2*Qx*Qy+Qy^2+20)

The first derivative with respect to Qx = (-6)*Qx+40-2*QY, the second
derivative with refer to Qx = -6, which is less than 0.

The first derivative with respect to QY = (-4)*Qy+30-2*Qx, the second
derivative with refer to QY = -4, which is less than 0.

Thus, the profit is maxium when the two equations are equal to 0.

Then set the two equations to be equal to 0,

and solve the two functions, we get Qx=7, and QY=4.

Therefore, Px=36, PY=26, and the maximum profit =

```{r}
7*(-2*7+50)+4*(-4+30)-(7^2+2*7*4+4^2+20)
```

7.  A manufacturer of parts for the tricycle industry sells three tires
    $(x)$ for every frame $(y)$. Thus, $\frac{x}{3}=y$ or $x=3 y$. If
    the demand is $$
    \begin{aligned}
    &x=63-0.25 P_{x} \\
    &y=60-\frac{1}{3} P_{y}
    \end{aligned}
    $$ and costs are $$
    c=x^{2}+x y+y^{2}+190
    $$ find the profit-maximizing level of (a) output, (b) price, and
    (c) profit.

Answer profit=Px*x+Py*y-cost
=x*((63-x)/0.25)+y*((60-y)*3)-(x\^2+x*y+y\^2+190)

Since x=3y,
profit=3*y*((63-3*y)/0.25)+y*((60-y)*3)-((3*y)^2+(3*y)*y+y^2+190)

The first derivative with respect to y = -104\*y+816

Then set the equation to be equal to 0, we get y=7.846154, and x =
23.53846.

Since the second derivative with respect to y(=-104) is less than 0,
this set of x and y values ensures the maximum profit.

We also get Px = 157.8462, Py=156.4615, and profit = 3952.769.

8.  Faced with two demand functions in two distinct markets $$
    Q_{1}=24-0.2 P_{1} \text { and } Q_{2}=10-0.05 P_{2}
    $$ and a cost function where the relationship between costs and
    quantity sold in either market is $T C=35+40 Q$, what prices will
    the firm charge (a) with price discrimination and (b) without price
    discrimination?

Answer

With price discrimination, the profit in market 1 =
P1*(24-0.2*P1)-(35+40*(24-0.2*P1)) = -0.2P1\^2+32P1-35-40\*24 take the
first derivative, and set it to 0, P1 = 80.

With price discrimination, the profit in market 1 =
P2*(10-0.05*P2)-(35+40*(10-0.05*P2)) = -0.05\*P2\^2+12P2-435 take the
first derivative, and set it to 0, P2 = 120.

Without price discrimination,P1 should be equal to P2, which is denoted
as P. the total profit in market 1 and 2 = P*Q1 + P*Q2 - TC1 - TC2 =
P*(24-0.2*P) + P*(10-0.05*P)-35-40*(24-0.2*P)-35-40*(10-0.05*P)

take the first derivative, and set it to 0, P = 88.

9.  A firm's sales grew from $\$ 1,000$ to $\$ 1,500$ from year one to
    year six. Express sales as an exponential function of time and
    determine the nominal rate of growth.

Answer

Since the sales of year 0 is 0, I set the exponential function as Sales
= x1(1+r)\^(t-1), where r means rate of growth.

According to the prompt, 1500=1000(1+r)\^5, thus r = 0.719.

10. Find the price and quantity that will maximize total revenue, given
    the demand function $P=12.50 e^{-0.005} Q$. Be sure to check the
    second-order condition.

Answer

Revenue = P*Q = P*((ln(P/12.5))/(-0.005)) take the first derivative,

```{r}
S=expression(P*((log(P/12.5))/(-0.005)))
```

```{r}
D(S, 'P')
```

=(log(P/12.5))/(-0.005))-200

take the second derivative,

```{r}
D(D(S, 'P'),'P')
```

```{r}
N=expression((log(P/12.5))/(-0.005)-200)
```

```{r}
D(N, 'P')
```

=-200/P

Since the second derivative is less than 0, the revenue is the maximum
when the first dirivative is equal to 0, and then P=4.598493, and Q=200.

```{r}
12.5*(exp(-1))
```

# Assignment 1

Question 1

Establish the shapes of the following sales response functions relating
sales Y in units per period to advertising expenditure X (\$ per period)

(i) Below function with a \> 0, b \> 0; Draw its graph. [5 points]

$$
Y=a+b \ln X
$$ Answer

```{r}
x=seq(0,100,0.01)
y=3 + 5*(log(x))
plot(x,y, type="l")
```

(ii) Below function with parameters a \> 0, b \> 0. Draw its graph [5
     points]

$$
Y=a e^{b X} \text { where } \mathrm{X}>0
$$ Answer

```{r}
x=seq(0,100,0.01)
y=5*((exp)(3*x))
plot(x,y, type="l")
```

(iii) Below function with parameters a \> 0, b \> 0 and c \>0. Draw its
      graph. [5 points]

$$
Y=\frac{a}{1+e^{-(b+c X)}}+d
$$ Answer

```{r}
x=seq(0,100,0.01)
y=5/((exp)(-(6+3*x))+1) + 3
plot(x,y, type="l")
```

Question 2

Consider Little's (short-term response) ADBUDG function that links
advertising efforts (A) to sales (S):

$$
S=S_0+S_1 \frac{A^\gamma}{\left(\rho+A^\gamma\right)}
$$

Explain the notation, i.e., name the variables and parameters of this
model (with their signs). Demonstrate the conditions when this function
is concave? Let m denote a constant gross constant margin. What is the
optimum advertising level when γ = 1? [10 points]

Answer

S denotes sales A denotes advertising efforts S0 denotes sales at t0 S1
denotes sales at t2

```{r}
S=expression(S0 + (s1 * A^gamma)/(rho+A^gamma))
```

The first derivative

```{r}
d1 = D(S, 'A')
d1
```

The simplified form

$$
\frac{\rho s_1 \gamma A^{\gamma-1}}{\left(A^\gamma+\rho\right)^2}
$$

The second derivative

```{r}
d2 = D(D(S,'A'),'A')
d2
```

The simplified form

$$
-\frac{\rho s_1 \gamma A^{\gamma-2} \cdot\left((\gamma+1) A^\gamma-\rho \cdot(\gamma-1)\right)}{\left(A^\gamma+\rho\right)^3}
$$

If f is twice-differentiable, then f is concave if and only if f ′′ is
non-positive (or, informally, if the "acceleration" is non-positive). If
its second derivative is negative then it is strictly concave, but the
converse is not true, as shown by f(x) = −x4.

The model is S-shaped for γ\>1 and concave for 0\<γ\<1.

When γ = 1, the optimum advertising level can be any amount.

Question 3. In the Tull et al (1986) paper, derive

(i) Equations (2) (optimum budget) and Equation (3) (ad elasticity) for
    the diminishing returns sales response function shown in Equation
    (1). (4 points) What are the sales at the optimum advertising level
    ? What is the maximized profit at the optimum advertising level?
    What is the elasticity at the optimum advertising level? (10 points)

Answer

Equations (1)

$$
Q=\alpha+\beta A^x
$$

\$\$

S = m\*(\alpha+\beta A\^x)

\$\$

```{r}
S=expression(m*(alpha+beta*A^x))
```

The first derivative

```{r}
d1 = D(S, 'A')
d1
```

The simplified form

$$
\beta m x A^{x-1}
$$ set the frist derivative equal to 0, how did they get A\* like this?

Equations (2)

$$
A^*=(x m \beta)^{1 / 1-x}
$$

Elasticity

Equations (3)

$$
E_{\mathrm{A}}=x\left(1-\frac{\alpha}{Q}\right)
$$

Calculate the maximized profit and the elasticity at the optimum
advertising level: plug in A\*

(iii) Equations (18) (optimum budget) and Equation (19) (ad elasticity)
      for the quadratic sales response function shown in Equation (17).
      What are the sales at the optimum advertising level? What is the
      maximized profit at the optimum advertising level? What is the
      elasticity at the optimum advertising level? (10 points)

Answer

Equations (17)

$$
Q=\alpha+\kappa A-\lambda A^2
$$

\$\$

S = m\*(\alpha+\kappa A-\lambda A\^2)

\$\$

```{r}
S=expression(m*(alpha+kappa*A-lambda*A^2))
```

The first derivative

```{r}
d1 = D(S, 'A')
d1
```

set the first derivative equal to 0: m \* (kappa - lambda \* (2 \* A)) =
0 A = kappa/2lambda

why did they get this?

Equations (18)

$$
A^*=\frac{m \kappa-1}{2 m \lambda}
$$

Equations (19)

$$
E_A *=\frac{\kappa-2 \lambda A *}{\kappa-\lambda A *}
$$

the sales at the optimum advertising level:

the maximized profit at the optimum advertising level:

the elasticity at the optimum advertising level:

4.  Qualitatively, how does the Mantrala et al (1992) paper modify and
    extend the managerial insights provided by Tull et al (1986) with
    regard to the 'flat maximum principle' in advertising decision
    making? (This is a short essay type question that should be brief --
    perhaps a couple of paragraphs - in less than a page.) (5 points)

Tull et al (1986) examined the impact of excessive advertising on
aggregate profits and sales under the assumptions of three types of
response functions: diminishing returns, saturation, and quadratic
sales. According to Tull et al (1986), although excessive advertising
can compromise firms' profits, the effect is subtle; Conversely,
excessive advertising can increase sales. In other words, if a firm's
goal is to increase market share, the firm can keep investing in
advertising to more than the optimal level at the subtle cost of profit
loss. It is called "flat maximum principle".

Mantrala et al (1992) extended the model by adding marketing resource
allocations to submarkets, such as geographic regions, market segments,
and product segments. They also assume alternative sales response
functions: concave saturation sales, S-shaped saturation sales, and
quadratic sales. According to Mantrala et al (1992), investors have to
consider investments and returns in the aggregate market and the
proportions of investments and returns in the submarkets. They argue
that Tull et al (1986) assumes the optimal resource applications in the
submarkets, however, if the allocation is less optimal and the response
function is concave saturation sales, the profits vary a lot, and the
investment in advertising may increase tremendously to reach the same
level of sales comparing to the optimal allocation level. In the
S-shaped function, resource allocation can even change the parameters
and the shape of the response function of the aggregate market. Further,
they account for investors' risk aversion levels, demonstrate different
optimal investment choices by risk-averse and risk-neutral investors,
and indicate profit variations.
